---
title: "   "
status: 
sample: 
output:
  html_document:
    code_folding: hide
    code_download: true
---
# Introduction

Research Question: 

Hypothesis:

# Initializion Code

```{r, message=FALSE}
# The user defines different variables and parameters that will be specific to the study.

injection_date <- as.Date("   ", format = "%m/%d/%Y")

trt_date <- as.Date("   ", format = "%m/%d/%Y")

sac_date <- as.Date("   ", format = "%m/%d/%Y")

engraftment_imaging_date <- as.Date("   ", format = "%m/%d/%Y")

imaging_dates <- as.Date(c("   ", "    ", "    "), format = "%m/%d/%Y")

number_of_expansion_mice <- 
number_of_groups <- 4
mice_per_group <- 
labels <- c("1" = "AAA", "2" = "BBB", "3" = "CCC", "4" = "DDD")
filtered_mice <- c(21:24, 26:40, 43)

# Load packages
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)

# Set the working directory to the current script's directory
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
```

# Data Processing

```{r, message=FALSE, warnings=FALSE}
# Create 'processed_data' which will be one data frame that contains:
# 1. all of the imaging data across weeks by funneling all of the weeks' imaging data into one data frame.
# 2. all of the mass data across weeks.

# Read the 'raw_data.csv' file into a data frame
raw_data <- read.csv("data/raw/raw_data.csv")

# Set the file path
file_path <- "data/imaging/processed/"

# Get a list of all weeks' imaging data .csv files in the directory
csv_files <- list.files(file_path, pattern = "\\.csv$", full.names = TRUE)

# Function to read, process, and save each csv file as a data frame with the same name
read_process_save_csv <- function(file) {
  # Extract the file name without extension
  file_name <- tools::file_path_sans_ext(basename(file))
  
  # Read the csv file into a data frame with check.names = FALSE
  df <- read.csv(file, check.names = FALSE)

# Remove specified columns
columns_to_remove <- c("Image Number", "ROI", "Image Layer", "Stdev Radiance", "Min Radiance", "Max Radiance")
df <- df[, !colnames(df) %in% columns_to_remove]
  
# Rename columns
colnames(df)[colnames(df) == "ROI Label"] <- "mouse_number"
colnames(df)[colnames(df) == "Experiment"] <- "imaging_date"
colnames(df)[colnames(df) == "Total Flux [p/s]"] <- "total_flux"

# Rename the last column header indirectly because otherwise, the exponent cannot be handled
  colnames(df)[ncol(df)] <- "avg_radiance"
  
  # Assign the processed data frame to a variable with the same name as the csv file
  assign(file_name, df, envir = .GlobalEnv)
}

# Use lapply() to read, process, and save each csv file
invisible(lapply(csv_files, read_process_save_csv))

# Combine all weeks' imaging data using do.call() and bind_rows()
invisible(imaging_data <- do.call(bind_rows, mget(tools::file_path_sans_ext(basename(csv_files)), envir = .GlobalEnv)))

# Convert imaging_date to Date class with the appropriate format
imaging_data$imaging_date <- as.Date(imaging_data$imaging_date, format = "%m/%d/%Y")

# Arrange the imaging_data data frame by imaging_date and then mouse_number
imaging_data <- imaging_data %>%
  arrange(imaging_date, mouse_number)
  
# Write the 'imaging_data' data frame to a CSV file
write_csv(imaging_data, file = "data/processed/imaging_data.csv")

# Combine 'imaging_data' with 'raw_data' to make processed_data
processed_data <- left_join(raw_data, imaging_data, by = "mouse_number")

# Read the 'raw_mass.csv' file into a data frame
raw_mass <- read.csv("data/raw/raw_mass.csv")

# Combine 'raw_mass' with 'processed_data'
processed_data <- left_join(processed_data, raw_mass, by = "mouse_number")

# Write the 'processed_data' data frame to a CSV file
write_csv(processed_data, file = "data/processed/processed_data.csv")
```

# Assignment {.tabset}

## Engraftment Status

```{r, message=FALSE, warnings=FALSE, fig.height=6, fig.width=11}
processed_data <- read.csv("data/processed/processed_data.csv")

processed_data %>% 
  ggplot(aes(x = reorder(mouse_number, -total_flux), y = total_flux, color = factor(imaging_date))) +
  geom_point() +
  scale_y_log10() +  # Apply logarithmic scale to the y-axis
  labs(x = "Mouse Number", y = "Total Flux", color = "Imaging Date") +
  theme_minimal()

# Count the number of mice with total_flux > 1e6 and total_flux <= 1e6
mice_greater_than_1e6 <- processed_data %>%
  filter(imaging_date == engraftment_imaging_date) %>% 
  distinct(trt, mouse_number, .keep_all = TRUE) %>%
  filter(total_flux > 1e6) %>%
  nrow()

mice_less_than_1e6 <- processed_data %>%
  filter(imaging_date == engraftment_imaging_date) %>% 
  distinct(trt, mouse_number, .keep_all = TRUE) %>%
  filter(total_flux <= 1e6) %>%
  nrow()

# Get the specific mouse numbers for mice with total_flux <= 1e6
specific_mice_numbers <- processed_data %>% filter(total_flux <= 1e6) %>% select(mouse_number)

# Create a table
table_data <- data.frame(
  "Engrafted" = mice_greater_than_1e6,
  "Not Engrafted" = mice_less_than_1e6
)

# Print the table
kable(table_data, format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Find the mouse numbers that are not engrafted, check if there are non-engrafted mice, and return the numbers
processed_data %>%
  filter(total_flux <= 1e6, imaging_date == engraftment_imaging_date) %>%
  pull(mouse_number) %>%
  unique() %>%
  { if (length(.) > 0) {
      cat("These mice are not engrafted because their total flux is less than 1e6:", paste(., collapse = ", "))
    } else {
      cat("All mice are engrafted because their total flux is greater than 1e6.")
    }
  }
```

## Treatment Assignment

Randomly assign mice to treatment conditions so that each treatment condition has equivalent total_flux between treatments. All of the unassigned remaining mice have the lowest total_flux and are assigned to the expansion group.

```{r, message=FALSE, warnings=FALSE}
filtered_data <- processed_data %>%
  arrange(total_flux) %>%
  filter(imaging_date == engraftment_imaging_date) %>%
  head(-number_of_expansion_mice)

# Define the assign_groups function
assign_groups <- function(data, num_groups, num_per_group) {
  # Create an empty vector to store group assignments
  trt <- integer(nrow(data))
  
  # Create a data frame to store mean total_flux for each group
  mean_total_flux <- data.frame(Group = 1:num_groups, Mean_Total_Flux = numeric(num_groups))
  
  # Shuffle the data to randomize assignments
  data <- data[sample(nrow(data)), ]
  for (i in 1:num_groups) {
    # Select a subset of data for each group
    group_data <- data[((i - 1) * num_per_group + 1):(i * num_per_group), ]
    
    # Calculate mean total_flux for the group
    mean_flux <- mean(group_data$total_flux)
    
    # Assign the group number to the trt column
    trt[((i - 1) * num_per_group + 1):(i * num_per_group)] <- i
    
    # Store the mean total_flux in the mean_total_flux data frame
    mean_total_flux[i, 2] <- mean_flux
  }
  
  # Add the trt column to the data frame
  data$trt <- trt
  
  # Return the data frame with group assignments
  return(list(Data = data, Mean_Total_Flux = mean_total_flux))
}

# Initialize variables to store the best seed and its corresponding variability
best_seed <- NULL
best_variability <- Inf

# Loop through the first 5000 seeds
for (seed in 1:5000) {
  set.seed(seed)
  
  # Call the function to assign groups using filtered_data
  result <- assign_groups(filtered_data, number_of_groups, mice_per_group)
  
  # Calculate the standard deviation of mean total_flux across groups
  variability <- sd(result$Mean_Total_Flux$Mean_Total_Flux)
  
  # Check if this seed has lower variability than the current best
  if (variability < best_variability) {
    best_variability <- variability
    best_seed <- seed
  }
}

# Print the seed with the least variability
cat("Seed with the least variability:", best_seed, "\n")

set.seed(best_seed)

# Function to assign mice to groups
assign_groups <- function(data, num_groups, num_per_group) {
  # Create an empty vector to store group assignments
  trt <- integer(nrow(data))
  
  # Create a data frame to store mean total_flux for each group
  mean_total_flux <- data.frame(Group = 1:num_groups, Mean_Total_Flux = numeric(num_groups))
  
  # Shuffle the data to randomize assignments
  data <- data[sample(nrow(data)), ]
  for (i in 1:num_groups) {
    # Select a subset of data for each group
    group_data <- data[((i - 1) * num_per_group + 1):(i * num_per_group), ]
    
    # Calculate mean total_flux for the group
    mean_flux <- mean(group_data$total_flux)
    
    # Assign the group number to the trt column
    trt[((i - 1) * num_per_group + 1):(i * num_per_group)] <- i
    
    # Store the mean total_flux in the mean_total_flux data frame
    mean_total_flux[i, 2] <- mean_flux
  }
  
  # Add the trt column to the data frame
  data$trt <- trt
  
  # Return the data frame with group assignments
  return(list(Data = data, Mean_Total_Flux = mean_total_flux))
}

# Call the function to assign groups using filtered_data
result <- assign_groups(filtered_data, number_of_groups, mice_per_group)

# Print the mean total_flux for each group
print(result$Mean_Total_Flux)

# Updated data frame with group assignments in the 'trt' column
filtered_data_1 <- result$Data

# Calculate the mean total_flux for each trt group
trt_means <- filtered_data %>%
  group_by(trt) %>%
  summarise(mean_total_flux = mean(total_flux))
  
# Calculate the mean of each trt group's total_flux
trt_means <- filtered_data %>%
  group_by(trt) %>%
  summarise(mean_total_flux = mean(total_flux))

# Add jitter to individual points to prevent overlap
jittered_data <- filtered_data %>%
  mutate(jittered_total_flux = total_flux + runif(n(), -0.2, 0.2))

# Create a scatter plot of means and individual points
ggplot() +
  geom_point(data = jittered_data, aes(x = trt, y = jittered_total_flux), 
             size = 2, color = "lightgray", alpha = 0.6, position = position_jitter(width = 0.2)) +
  geom_point(data = trt_means, aes(x = trt, y = mean_total_flux), 
             size = 3, color = "blue") +
  labs(x = "trt", y = "Total Flux") +
  theme_minimal()

# Merge the data frames filtered_data and processed_data
processed_data <- processed_data %>%
  left_join(filtered_data %>% select(mouse_number, trt), by = "mouse_number") %>%
  mutate(trt = coalesce(trt.x, trt.y)) %>%
  select(-trt.x, -trt.y) %>%
  select(mouse_number, trt, everything()) %>%
  relocate(trt, .before = trt_injection_vial)
  
# Write the 'processed_data' data frame to a CSV file
write_csv(processed_data, file = "data/processed/processed_data.csv")
```

# Procedures {.tabset}

## Thawing tumor:

```{r, message=FALSE, warnings=FALSE}
processed_data <- read.csv("data/processed/processed_data.csv")

processed_data %>%
  select(mouse_number, tumor_injection_quality, tumor_injection_vial, tumor_cells_per_mouse, tumor_injection_date, tumor_total_cell_count, tumor_cell_viability_after_thaw, tumor_total_live_cell_count) %>% 
  filter(mouse_number %in% filtered_mice) %>% 
  distinct(mouse_number, .keep_all = TRUE) %>%
  kable(format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Thawing trt:

```{r, message=FALSE, warnings=FALSE}
processed_data <- read.csv("data/processed/processed_data.csv")

processed_data %>%
  select(mouse_number, trt_injection_quality, trt_injection_vial, trt_cells_per_mouse, trt_injection_date, trt_total_cell_count, trt_cell_viability_after_thaw, trt_total_live_cell_count) %>% 
  filter(mouse_number %in% filtered_mice) %>% 
  distinct(mouse_number, .keep_all = TRUE) %>%
  kable(format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

# Subjects {.tabset}

## Mouse Current Status Plot

```{r, message=FALSE, warnings=FALSE, fig.height=6, fig.width=10}
processed_data <- read.csv("data/processed/processed_data.csv")

# Your summary table
summary_table <- processed_data %>%
  filter(mouse_number %in% filtered_mice) %>%
  distinct(trt, mouse_number, .keep_all = TRUE) %>%
  group_by(trt) %>%
  summarize(
    Dead = n_distinct(mouse_number[!is.na(manner_of_death)]),  # Count unique mice with non-missing values for 'Dead'
    Alive = n_distinct(mouse_number[is.na(manner_of_death)]),   # Count unique mice with missing values for 'Alive'
    Dead_Mouse_Numbers = toString(mouse_number[!is.na(manner_of_death)]),  # List mouse numbers for Dead
    Alive_Mouse_Numbers = toString(mouse_number[is.na(manner_of_death)])  # List mouse numbers for Alive
  )

# Calculate total unique mice dead and alive
total_dead <- sum(summary_table$Dead)
total_alive <- sum(summary_table$Alive)

# Create a grouped bar plot
p <- ggplot(summary_table, aes(x = factor(trt, labels = labels))) +
  geom_bar(aes(y = Dead, fill = "Dead"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = Alive, fill = "Alive"), stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Dead" = "red", "Alive" = "green")) +
  labs(x = "Treatment", y = "Count", fill = "Status") +
  ggtitle("Dead and Alive Mice Counts by Treatment") +
  theme_minimal() +
  theme(legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())  # Remove horizontal grid lines between y-axis tick marks

# Set x-axis breaks and limits using scale_x_discrete
p <- p + scale_x_discrete(breaks = labels, limits = labels)

# Set y-axis breaks and limits
p <- p +
  scale_y_continuous(
    breaks = seq(0, 5, by = 1),  # Set breaks by units of 1
    limits = c(0, 8)  # Adjust the limits to provide more space for labels
  )

# Add annotations for alive and dead mouse numbers with color based on 'death_date'
p <- p +
  geom_text(aes(x = factor(trt, labels = labels), y = Dead + Alive + 0.5, 
                label = Dead_Mouse_Numbers, color = "red"), 
            size = 5, hjust = 0.5) +  # Adjust the size parameter for larger labels
  geom_text(aes(x = factor(trt, labels = labels), y = Dead + Alive + 0.5, 
                label = Alive_Mouse_Numbers, color = "green"), 
            size = 5, hjust = 0.5)  # Adjust the size parameter for larger labels

# Set legend labels including total counts
p <- p +
  labs(fill = paste("Status (Alive:", total_alive, ", Dead:", total_dead, ")"))

# Specify the color scale for text labels
p <- p +
  scale_color_manual(values = c("red" = "red", "green" = "green"))

# Remove the color legend for "Status"
p <- p +
  guides(color = "none")

print(p)

# Convert date columns to date format
processed_data$imaging_date <- as.Date(processed_data$imaging_date)
processed_data$mass_date <- as.Date(processed_data$mass_date)
processed_data$death_date <- as.Date(processed_data$death_date)

# Find the most recent imaging_date
most_recent_imaging_date <- max(processed_data$imaging_date, na.rm = TRUE)

# Find the oldest and newest mass_date for each mouse
mass_date_summary <- processed_data %>%
  filter(!is.na(mass_date)) %>%
  group_by(mouse_number) %>%
  summarize(
    oldest_mass_date = min(mass_date),
    newest_mass_date = max(mass_date)
  )

# Calculate the mass loss percentage for each mouse
mass_loss <- mass_date_summary %>%
  mutate(
    mouse_number = as.character(mouse_number),  # Convert mouse_number to character
    mass_loss_percentage = (1 - newest_mass_date / oldest_mass_date) * 100
  )

# Find mice with total_flux >= 2e10 from the most recent imaging_date
high_flux_mice <- processed_data %>%
  filter(imaging_date == most_recent_imaging_date,
         total_flux >= 2e10,
         is.na(death_date))

# Find mice meeting the mass loss criterion (20% or more loss)
mass_loss_mice <- mass_loss %>%
  filter(mass_loss_percentage >= 20)

# Combine the two sets of mice that meet either criteria
mice_to_euthanize <- union(high_flux_mice$mouse_number, mass_loss_mice$mouse_number)

# Check if there are any mice to euthanize based on either criterion
if (length(mice_to_euthanize) > 0) {
  if (length(mice_to_euthanize) > 0 && length(high_flux_mice$mouse_number) > 0) {
    cat("With respect to the IACUC protocol, you need to euthanize mouse number(s) meeting the following criteria:\n")
    cat("- Total flux greater than or equal to 2x10^10:", paste(high_flux_mice$mouse_number, collapse = ", "), "\n")
  }
  
  if (length(mice_to_euthanize) > 0 && length(mass_loss_mice$mouse_number) > 0) {
    cat("- Mass loss of 20% or more:", paste(mass_loss_mice$mouse_number, collapse = ", "), "\n")
  }
} else {
  cat("No mice meet the criteria for euthanasia.")
}

# Create a scatterplot for cage_number, trt, and status without overlap
scatterplot <- processed_data %>%
  select(cage_number, mouse_number, trt, manner_of_death) %>% 
  filter(mouse_number %in% filtered_mice) %>% 
  distinct(trt, mouse_number, .keep_all = TRUE) %>% 
  ggplot(aes(x = factor(cage_number), y = factor(trt, labels = labels))) +  # Use labels here
  geom_jitter(
    aes(color = ifelse(is.na(manner_of_death) | manner_of_death == "", "Alive", "Dead")),
    size = 3,
    position = position_dodge2(width = 0.5),  # Adjust width as needed
    alpha = 0.7
  ) +
  geom_text(
    aes(label = mouse_number), 
    position = position_dodge2(width = 0.5),  # Match the jitter width
    vjust = -1,  # Adjust vertical position to add space
    size = 3
  ) +
  scale_color_manual(
    values = c("Dead" = "red", "Alive" = "green")
  ) +
  labs(x = "Cage Number", y = "Treatment", color = "Status") +
  ggtitle("Mouse Status by Cage and Treatment") +
  theme_minimal() +
  theme(legend.position = "top", 
        axis.text.x = element_text(size = 12, margin = margin(0, 40, 0, 40))) +  # Remove angle and adjust size
  guides(color = "none") +  # Remove the legend for "Status"
  scale_x_discrete(breaks = unique(raw_data$cage_number))  # Set breaks for x-axis

print(scatterplot)
```

## Mouse Current Status Table

```{r, message=FALSE, warnings=FALSE}
# Create and print the table with labels
table_data <- processed_data %>%
  select(mouse_number, death_date, manner_of_death) %>% 
  filter(mouse_number %in% filtered_mice) %>% 
  distinct(mouse_number, .keep_all = TRUE)

processed_data %>%
  select(cage_number, mouse_number, trt, imaging_date, total_flux, avg_radiance, manner_of_death) %>% 
  filter(mouse_number %in% filtered_mice) %>%
  mutate(
    trt = labels[as.character(trt)],
    log_total_flux = log(total_flux),
    days_from_inj_to_death = exp(-0.1728 * log_total_flux + 5.1164),
    predicted_death_date = as.Date(Sys.Date() + days_from_inj_to_death, origin = "1970-01-01")
  ) %>%
  group_by(mouse_number) %>%
  filter(imaging_date == max(imaging_date)) %>%
  kable(format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Create a .csv file of current mice data to print
table_data <- processed_data %>%
  select(cage_number, mouse_number, trt, imaging_date, total_flux, avg_radiance, death_date, manner_of_death) %>%
  filter(mouse_number %in% c(21:44), imaging_date == max(imaging_date)) %>% 
  arrange(cage_number, mouse_number, trt) %>%
  mutate(
    trt = labels[as.character(trt)],
    log_total_flux = log(total_flux),
    days_from_inj_to_death = exp(-0.1728 * log_total_flux + 5.1164),
    predicted_death_date = as.Date(Sys.Date() + days_from_inj_to_death, origin = "1970-01-01")
  ) %>%
  mutate(
    # Calculate standard error for the days_from_inj_to_death
    stderr = 1.96 * (days_from_inj_to_death / sqrt(n())),
    lower_bound = predicted_death_date - stderr,
    upper_bound = predicted_death_date + stderr
  ) %>%
  select(-log_total_flux, -days_from_inj_to_death, -stderr)

# Write the data to a CSV file named "current_mice_data.csv"
write.csv(table_data, file = "current_mice_data.csv", row.names = FALSE)
```

## Cage Cards

```{r, results='asis'}
png_files <- list.files("cage_cards", pattern = "\\.png$", full.names = TRUE)

for (png_file in png_files) {
  cat(sprintf("<img src='%s' />\n\n", png_file))
}
```

# Results {.tabset}

## Plots {.tabset}

### Mean Total Flux by Treatment

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=11}
processed_data <- read.csv("data/processed/processed_data.csv")

# Group by trt and imaging_date, and calculate the mean total_flux for each group
mean_data <- processed_data %>%
  filter(mouse_number %in% filtered_mice) %>% 
  group_by(trt, imaging_date) %>%
  summarise(mean_total_flux = mean(total_flux)) %>%
  mutate(imaging_date = as.Date(imaging_date))

# Define a custom theme for the plot (without vertical grid lines)
my_theme <- theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12, margin = margin(r = 15)),  # Adjust the margin
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    panel.grid.major.x = element_blank(),  # Remove vertical grid lines
    panel.grid.minor.x = element_blank()   # Remove vertical minor grid lines
  )

# Convert imaging_dates to character strings
imaging_date_breaks <- as.character(imaging_dates)

# Create the scatter plot with lines connecting group points
p <- ggplot(data = mean_data, aes(x = as.Date(imaging_date), y = mean_total_flux, color = factor(trt))) + 
  geom_point(position = position_dodge(width = 0.4), size = 1.3) +  
  geom_line(aes(group = factor(trt)), position = position_dodge(width = 0.4), size = 0.9) +
  labs(x = "Imaging Date", y = "Mean Total Flux", color = "Treatment") +
  my_theme +
  scale_y_log10(labels = scales::scientific_format()) +  # Make the y-axis a log scale
  geom_vline(xintercept = as.Date(c(injection_date, trt_date, sac_date), format = "%m/%d/%Y"), 
             linetype = "dashed", color = "black") +
  scale_color_manual(
    values = c(
      "1" = alpha("#f8766d", 1),  # XXX
      "2" = alpha("#7cae00", 1),  # XXX
      "3" = alpha("#00bfc4", 1),  # XXX
      "4" = alpha("#C3B1E1", 1)   # XXX
    ),
    labels = labels  # Assign custom labels to trt levels
  ) +
  scale_x_continuous(
    breaks = as.numeric(imaging_dates),  # Set custom breaks as numeric values
    labels = as.character(imaging_dates)  # Convert breaks to character labels
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  # Add labels for vertical lines
  geom_text(
    x = as.Date(injection_date, format = "%m/%d/%Y"),  # X position for the label
    y = Inf,  # Position the label outside the plot area
    label = "ETP1 Injection",  # Label text
    vjust = 1,  # Adjust vertical position
    hjust = -0.1,  # Adjust horizontal position
    color = "black"  # Set the text color to black
  ) +
  geom_text(
    x = as.Date(trt_date, format = "%m/%d/%Y"),  # X position for the label
    y = Inf,  # Position the label outside the plot area
    label = "Treatment",  # Label text
    vjust = 1,  # Adjust vertical position
    hjust = -0.1,  # Adjust horizontal position
    color = "black"  # Set the text color to black
  ) +
  geom_text(
    x = as.Date(sac_date, format = "%m/%d/%Y"),  # X position for the label
    y = Inf,  # Position the label outside the plot area
    label = "Sacrifice",  # Label text
    vjust = 1,  # Adjust vertical position
    hjust = 1.1,  # Adjust horizontal position to the left
    color = "black"  # Set the text color to black
  ) +
  theme(legend.text = element_text(size = 10))  # Adjust legend text size

print(p)
```

### Mean Total Flux and Individual Total Flux by Treatment

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=11}
# Read the data
processed_data <- read.csv("data/processed/processed_data.csv")

# Group by trt and imaging_date, and calculate the mean total_flux for each group
mean_data <- processed_data %>%
  filter(mouse_number %in% filtered_mice) %>% 
  group_by(trt, imaging_date) %>%
  summarise(mean_total_flux = mean(total_flux, na.rm = TRUE)) %>%
  mutate(imaging_date = as.Date(imaging_date))

# Define a custom theme for the plot (without vertical grid lines)
my_theme <- theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12, margin = margin(r = 15)),  # Adjust the margin
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    panel.grid.major.x = element_blank(),  # Remove vertical grid lines
    panel.grid.minor.x = element_blank(),  # Remove vertical minor grid lines
    strip.text.x = element_blank()  # Remove the titles above facets
  )

# Create the plot, including filtering for mouse_number
p <- ggplot(data = processed_data %>%
              filter(mouse_number %in% filtered_mice), 
            aes(x = as.Date(imaging_date), y = total_flux, color = factor(trt))) +
  geom_line(data = mean_data, aes(x = imaging_date, y = mean_total_flux, group = trt), size = 0.9) +  # Connected lines for group means
  geom_point(size = 1.3, shape = 16, alpha = 0.5) +  # Points for individual mice
  labs(x = NULL, y = "Total Flux", color = "Treatment") +
  scale_y_log10(labels = scales::scientific_format()) +
  scale_color_manual(
    values = c(
      "1" = alpha("#f8766d", 1),  # XXX
      "2" = alpha("#7cae00", 1),  # XXX
      "3" = alpha("#00bfc4", 1),  # XXX
      "4" = alpha("#C3B1E1", 1)   # XXX
    ),
    labels = labels
  ) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  my_theme +
  facet_wrap(~ trt)

# Print the plot with corrected aesthetics for mouse total_flux points and the legend to the right
print(p + theme(legend.position = "right", legend.direction = "vertical", legend.box = "vertical"))
```

### Total Flux of Individual Mice  by Treatment

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=11}
# Read the data
processed_data <- read.csv("data/processed/processed_data.csv")

# Define a custom theme for the plot (without vertical grid lines)
my_theme <- theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12, margin = margin(r = 15)),  # Adjust the margin
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    panel.grid.major.x = element_blank(),  # Remove vertical grid lines
    panel.grid.minor.x = element_blank(),  # Remove vertical minor grid lines
    strip.text.x = element_blank()  # Remove the titles above facets
  )

# Create the plot, including filtering for mouse_number
p <- ggplot(data = processed_data %>%
              filter(mouse_number %in% filtered_mice), 
            aes(x = as.Date(imaging_date), y = total_flux, color = factor(trt))) +
  geom_point(aes(group = interaction(trt, mouse_number)), size = 2.5, shape = 16, alpha = 0.5) +  # Thicker points for individual mice
  geom_line(aes(group = interaction(trt, mouse_number)), size = 1.2, alpha = 0.3) +  # Thicker lines connecting points for each mouse
  labs(x = NULL, y = "Total Flux", color = "Treatment") +
  scale_y_log10(labels = scales::scientific_format()) +
  scale_color_manual(
    values = c(
      "1" = alpha("#f8766d", 1),  # XXX
      "2" = alpha("#7cae00", 1),  # XXX
      "3" = alpha("#00bfc4", 1),  # XXX
      "4" = alpha("#C3B1E1", 1)   # XXX
    ),
    labels = labels
  ) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  my_theme +
  facet_wrap(~ trt)

# Print the plot with corrected aesthetics for mouse total_flux points and the legend to the right
print(p + theme(legend.position = "right", legend.direction = "vertical", legend.box = "vertical"))
```

### Median Total Flux by Treatment

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=11}
processed_data <- read.csv("data/processed/processed_data.csv")

# Group by trt and imaging_date, and calculate the median total_flux for each group
median_data <- processed_data %>%
  filter(mouse_number %in% filtered_mice) %>% 
  group_by(trt, imaging_date) %>%
  summarise(median_total_flux = median(total_flux)) %>%
  mutate(imaging_date = as.Date(imaging_date))

# Define a custom theme for the plot (without vertical grid lines)
my_theme <- theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12, margin = margin(r = 15)),  # Adjust the margin
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    panel.grid.major.x = element_blank(),  # Remove vertical grid lines
    panel.grid.minor.x = element_blank()   # Remove vertical minor grid lines
  )

# Convert imaging_dates to character strings
imaging_date_breaks <- as.character(imaging_dates)

# Create the scatter plot with lines connecting group points
p <- ggplot(data = median_data, aes(x = as.Date(imaging_date), y = median_total_flux, color = factor(trt))) + 
  geom_point(position = position_dodge(width = 0.4), size = 1.3) +  
  geom_line(aes(group = factor(trt)), position = position_dodge(width = 0.4), size = 0.9) +
  labs(x = "Imaging Date", y = "Median Total Flux", color = "Treatment") +
  my_theme +
  scale_y_log10(labels = scales::scientific_format()) +
  geom_vline(xintercept = as.Date(c(injection_date, trt_date, sac_date), format = "%m/%d/%Y"), 
             linetype = "dashed", color = "black") +
  scale_color_manual(
    values = c(
      "5" = alpha("#f8766d", 1),  # CCR4 CAR2
      "6" = alpha("#7cae00", 1),  # CART38
      "7" = alpha("#00bfc4", 1),  # UTD CART38
      "8" = alpha("#C3B1E1", 1)   # Untreated
    ),
    labels = labels  # Assign custom labels to trt levels
  ) +
  scale_x_continuous(
    breaks = as.numeric(imaging_dates),  # Set custom breaks as numeric values
    labels = as.character(imaging_dates)  # Convert breaks to character labels
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  # Add labels for vertical lines
  geom_text(
    x = as.Date(injection_date, format = "%m/%d/%Y"),  # X position for the label
    y = Inf,  # Position the label outside the plot area
    label = "ETP1 Injection",  # Label text
    vjust = 1,  # Adjust vertical position
    hjust = -0.1,  # Adjust horizontal position
    color = "black"  # Set the text color to black
  ) +
  geom_text(
    x = as.Date(trt_date, format = "%m/%d/%Y"),  # X position for the label
    y = Inf,  # Position the label outside the plot area
    label = "Treatment",  # Label text
    vjust = 1,  # Adjust vertical position
    hjust = -0.1,  # Adjust horizontal position
    color = "black"  # Set the text color to black
  ) +
  geom_text(
    x = as.Date(sac_date, format = "%m/%d/%Y"),  # X position for the label
    y = Inf,  # Position the label outside the plot area
    label = "Sacrifice",  # Label text
    vjust = 1,  # Adjust vertical position
    hjust = 1.1,  # Adjust horizontal position to the left
    color = "black"  # Set the text color to black
  ) +
  theme(legend.text = element_text(size = 10))  # Adjust legend text size

print(p)
```

### Median Total Flux and Individual Total Flux by Treatment

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=11}
# Read the data
processed_data <- read.csv("data/processed/processed_data.csv")

# Group by trt and imaging_date, and calculate the median total_flux for each group
median_data <- processed_data %>%
  filter(mouse_number %in% filtered_mice) %>% 
  group_by(trt, imaging_date) %>%
  summarise(median_total_flux = median(total_flux, na.rm = TRUE)) %>%
  mutate(imaging_date = as.Date(imaging_date))

# Define a custom theme for the plot (without vertical grid lines)
my_theme <- theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12, margin = margin(r = 15)),  # Adjust the margin
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    panel.grid.major.x = element_blank(),  # Remove vertical grid lines
    panel.grid.minor.x = element_blank(),  # Remove vertical minor grid lines
    strip.text.x = element_blank()  # Remove the titles above facets
  )

# Create the plot, including filtering for mouse_number
p <- ggplot(data = processed_data %>%
              filter(mouse_number %in% filtered_mice), 
            aes(x = as.Date(imaging_date), y = total_flux, color = factor(trt))) +
  geom_line(data = median_data, aes(x = imaging_date, y = median_total_flux, group = trt), size = 0.9) +  # Connected lines for group medians
  geom_point(size = 1.3, shape = 16, alpha = 0.5) +  # Points for individual mice
  labs(x = NULL, y = "Total Flux", color = "Treatment") +
  scale_y_log10(labels = scales::scientific_format()) +
  scale_color_manual(
    values = c(
      "5" = alpha("#f8766d", 1),  # CCR4 CAR2
      "6" = alpha("#7cae00", 1),  # CART38
      "7" = alpha("#00bfc4", 1),  # UTD CART38
      "8" = alpha("#C3B1E1", 1)   # Untreated
    ),
    labels = labels
  ) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  my_theme +
  facet_wrap(~ trt)

# Print the plot with corrected aesthetics for mouse total_flux points and the legend to the right
print(p + theme(legend.position = "right", legend.direction = "vertical", legend.box = "vertical"))
```

## Data Table

```{r, message=FALSE, warnings=FALSE}
processed_data %>%
  select(cage_number, mouse_number, trt, manner_of_death, imaging_date, total_flux, avg_radiance) %>% 
  filter(mouse_number %in% filtered_mice) %>% 
  filter(!is.na(trt)) %>%
  mutate(trt = labels[as.character(trt)]) %>% 
  kable(format = "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Kaplan-Meier Curve

```{r, message=FALSE, warning=FALSE, fig.height=10, fig.width=11}
processed_data <- read.csv("data/processed/processed_data.csv")

# Check if processed_data has any rows before proceeding
if (nrow(processed_data) > 0) {
  # Filter the data by mouse_number
  processed_data <- processed_data %>%
    filter(mouse_number %in% filtered_mice)
  
  # Convert the columns to Date format with the format 'month/day/year'
  processed_data <- processed_data %>%
    mutate(
      death_date = as.Date(death_date, format = "%m/%d/%Y"),
      trt_injection_date = as.Date(trt_injection_date, format = "%m/%d/%Y"),
      tumor_injection_date = as.Date(tumor_injection_date, format = "%m/%d/%Y")
    )
  
  # Calculate the days_from_trt_to_death using both trt_injection_date and tumor_injection_date
  processed_data <- processed_data %>%
    mutate(
      days_from_trt_to_death = as.numeric(difftime(death_date, trt_injection_date, units = "days")),
      # Replace NA values in days_from_trt_to_death with the corresponding difference using tumor_injection_date
      days_from_trt_to_death = ifelse(is.na(days_from_trt_to_death), as.numeric(difftime(death_date, tumor_injection_date, units = "days")), days_from_trt_to_death)
    )
  
  # Create a survival object
  survival_data <- with(processed_data, Surv(days_from_trt_to_death, event = rep(1, length(trt))))
  
  # Check if there are any non-missing observations in survival_data
  if (any(!is.na(survival_data))) {
    # Fit Kaplan-Meier survival curves for each treatment group
    surv_fit <- survfit(survival_data ~ trt, data = processed_data)
    
    # Plot Kaplan-Meier curves
    ggsurvplot(surv_fit, data = processed_data, risk.table = TRUE, pval = TRUE)
    
    # Check if the days_from_trt_to_death column exists before selecting it
    if ("days_from_trt_to_death" %in% colnames(processed_data)) {
      # Create a table that shows the rounded average days of survival by trt
      average_days_by_trt <- processed_data %>%
        group_by(trt) %>%
        summarize(average_survival_in_days = round(mean(days_from_trt_to_death, na.rm = TRUE)))
      
      # Create and print the table
      average_days_by_trt %>%
        kable("html") %>%
        kable_styling(bootstrap_options = "striped", full_width = FALSE)
    } else {
      cat("Column 'days_from_trt_to_death' not found in processed_data.")
    }
  } else {
    cat("No non-missing observations for survival data found.")
  }
} else {
  cat("No data available for analysis.")
}

# Select the columns you want to include in the table and filter by mouse_number
table_data <- raw_data %>%
  filter(mouse_number %in% filtered_mice) %>%
  select(
    cage_number,
    mouse_number,
    trt,
    manner_of_death
  )

# Create and print the table
kable(table_data, "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## ANOVA

Do not take this result seriously, still working on ANOVA + other descriptives info

```{r, message=FALSE, warning=FALSE}
# Fit an ANOVA model
anova_model <- aov(total_flux ~ trt, data = processed_data)

# Perform ANOVA analysis
anova_result <- summary(anova_model)

# Print the ANOVA results
print(anova_result)
```

## Spleen and Marrow

```{r, message=FALSE, warning=FALSE, fig.height=10, fig.width=15}
processed_data <- read.csv("data/processed/processed_data.csv")

bar_plot <- processed_data %>% 
  filter(mouse_number %in% filtered_mice) %>%
  mutate(
    cryovial_spleen_cell_total = coalesce(cryovial_spleen_cell_total, 0),
    cryovial_marrow_cell_total = coalesce(cryovial_marrow_cell_total, 0)
  ) %>%
  ggplot() +
  geom_bar(aes(x = mouse_number - 0.2, y = cryovial_spleen_cell_total, fill = "Spleen"), stat = "identity", position = "identity", width = 0.4) +
  geom_bar(aes(x = mouse_number + 0.2, y = cryovial_marrow_cell_total, fill = "Marrow"), stat = "identity", position = "identity", width = 0.4) +
  scale_fill_manual(values = c("Spleen" = "blue", "Marrow" = "red")) +
  labs(x = "Mouse Number", y = "Cell Total", fill = "Tissue") +
  ggtitle("Cryovial Spleen and Marrow Cell Total by Mouse Number") +
  theme_minimal() +
  theme(legend.position = "top") +
  scale_x_continuous(breaks = unique(processed_data$mouse_number), labels = unique(processed_data$mouse_number)) +
  coord_cartesian(ylim = c(0, max(processed_data$cryovial_spleen_cell_total, processed_data$cryovial_marrow_cell_total) + 10))

# Print the bar plot
print(bar_plot)

# Select the columns you want to include in the table
table_data <- processed_data %>%
  filter(mouse_number %in% filtered_mice) %>% 
  distinct(trt, mouse_number, .keep_all = TRUE) %>%
  select(
    cage_number,
    mouse_number,
    trt,
    manner_of_death,
    number_of_spleen_cryovials,
    cryovial_spleen_cell_total,
    number_of_marrow_cryovials,
    cryovial_marrow_cell_total
  ) %>% 
  mutate(trt = labels[as.character(trt)])

# Format columns to scientific notation
table_data <- table_data %>%
  mutate(
    cryovial_marrow_cell_total = format(cryovial_marrow_cell_total, scientific = TRUE)
  )

# Create and print the table
kable(table_data, "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```
